---
title: "P8160 Monte Carlo Simulation Design"
subtitle: "Compare three Survival models"
author: "Jingchen Chai, Yi Huang, Ruihan Zhang"
output:
  powerpoint_presentation:
    reference_doc: Mailman_Standard_Template.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# install.packages('xaringan')
if (packageVersion('rmarkdown') <= '1.9') {
  install.packages('rmarkdown')  # update rmarkdown from CRAN
}
library(knitr)
library(kableExtra)
```
## Background


## Methodology
Survival function measures the probability of observing individual survival time $T$ beyond a certain time $t$\
$S(t) = Pr(T>t) = \int_t^{\infty}f(s) ds$\
Cumulative Distribution Function of survival time T\
$F(t) = Pr(T \leq t) = 1 - S(t) = 1 - \int_t^{\infty}f(s) ds$\
Hazard Function\
$h(t) = \displaystyle{\lim_{\Delta_t \to 0}}\frac{Pr((T\in(t,t+\Delta_t)|T>t)}{\Delta_t} = \frac{f(t)}{S(t)}$ \


## ...
Proportional hazard model is the primary regression model to investigate the effectiveness of treatment $X$ over survival time T, where the i-th patient at a time t \
is\
$h_i(t) = h_0(t)e^{x_i\beta}$\
where\
- $h_0$ the baseline hazard function\
- $x_i$ is the treatment indicator variable (control = 0, treatment = 1)\
- $\beta$ is the log hazard ratio of treatment effect\

Cumulative Hazard Function\
$H(t) = \int_t^{\infty}h(t)dt = \int_t^{\infty}h_0(t)e^{X\beta}dt = H_0(t)e^{X\beta}$\
where\
- $H_0(t)$ is the cumulative baseline risk\


 



## ...

The proportional hazard can be expressed as \
\
$\frac{h(t|x_0)}{h(t|x_1)} = e^{\beta({x_0-x_1})}$\

Derive survival time T \
\
$H(t)=-\log{(S(t))}$\
$S(t)=e^{-H(t)}=e^{\int_t^\infty h(s)ds}$\

By Inverse Transformation Method\

$T = F^{-1}(U) = H_0^{-1}(\frac{-\log{(U)}}{e^{X\beta}})$\
where $U\sim U(0,1)$\

## ...

$T = F^{-1}(U) = H_0^{-1}(\frac{-\log{(U)}}{e^{X\beta}})$\
where $U\sim U(0,1)$\
\
In our simulation, we will consider three types of data: \
Exponential, Weibull, and Gompertz\

Apply Inverse transformation method, obtain the following table\

## ...
**Generate data**\
1. Define Random Variable $X$, treatment, from a binomial distribution with $p=0.5$\
2. Generate survival time $T$, time to event, using $X$ and $\beta$\
3. Randomly generate censoring time $C$, from an exponential distribution\
4. We observe either the survival time T or else the censoring time C. \
Specifically, we observe the random variable\
$Y =min(T,C)$\
Create a status indicator variable (1 = event, 0 = censored)\
$$
Status = 
    \begin{cases}
      1, & T_{i}\le C_{i} \\
      0, &  T_{i} > C_{i}
    \end{cases}
$$
\
**Define true treatment effect and parameter ** \
1. True treatment effect $\beta = 2$\
2. 7 different sample size N ranging from 100 to 400 increasing by 50 \
3. Simulate exponential distribution with $\lambda=0.5$\
4. Simulate weibull distribution with $\lambda=0.5$ and $\gamma=2$\
5. Simulate gompertz distribution with $\lambda=0.5$ and $\alpha=2$\


## ...
$\hat{\beta}$\
\
$Bias = \frac{1}{k}\sum_{i=1}^k(\hat{\beta^{(k)}}-\beta)=\frac{1}{k}\sum_{i=1}^k(\hat{\beta^{(k)}}-2)$\
$Variance = \frac{1}{k-1}\sum_{i=1}^k\hat{(\beta^{(k)}}-\beta)^2= \frac{1}{k-1}\sum_{i=1}^k\hat{(\beta^{(k)}}-2)^2$\
$MSE = \frac{1}{k}\sum_{i=1}^k(\hat{\beta^{(k)}}-\beta)^2 = \frac{1}{k}\sum_{i=1}^k(\hat{\beta^{(k)}}-2)^2$\

## ...
Survival function, $S(t)$ is the probability of observing individual survival time T beyond a certain time $t$.\

The Cumulative distribution function of the random variable survival time T, F(t)=Pr(T$\leq$t) is the probability of observing individual survival time T less than a certain time $t$. \

The Hazard function, $h(t)$ is the product of $h_0(t)$ and $r(X, \beta)$, which is defined as the ratio of probability density function of time variable and Survival function\

where $h_0(t)$ is the baseline hazard function, how the hazard function changes as a function of survival time T\
$r(X, \beta)=exp(X\beta)$ is a function characterize how the hazard function changes as a function of covariate $X_{treatment}$\

Cumulative hazard function i



## R Markdown
This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks w ithin the document.

## Slide with Results

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

