---
title: "P8160_Project1_Survival"
author: "Yi Huang"
date: "2023-02-19"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.align = "center", warning = FALSE)
library(simsurv)
library(survival)
library(tidyverse)
library(ggplot2)
# install.packages("tidyr")
library(tidyr)
# install.packages("flexsurv")
library(flexsurv)
```

# Project 1: Design a simulation study to compare three survival models

# Plot of Baseline Hazard Function
```{r}
# parameter and seed
set.seed(2023)

# gamma = 2, alpha = 2
exp_haz <- function(t, lambda = 0.5) {lambda * 1 * t^0}
weibull_haz_2 <- function(t, lambda = 0.5) lambda * 2 * t^(2 - 1)
gompertz_haz_2 <- function(t, lambda = 0.5) lambda * exp(2* t)

p1 <- ggplot(data.frame(x=c(0, 5)), aes(x=x)) + 
    geom_path(aes(colour="red"), stat="function", fun=exp_haz)+
    geom_path(aes(colour="green"), stat="function", fun=weibull_haz_2)+
    geom_path(aes(colour="blue"), stat="function", fun=gompertz_haz_2) +
    scale_colour_identity("Hazard Function", guide="legend", 
                          labels = c("Exponential", "Weibull(gamma = 2)", "Gompertz (alpha = 2)"), 
                          breaks = c("red", "green", "blue")) +
    labs(x = "t",
         y = "h0(t)") + ylim(0,2)
p1
```




# Simulation for Gompertz Baseline Hazard Function

- N: sample size 100, 150, 200, 250, 300, 350, 400
- m: simulation time 1000
- $\beta$: true treatment effect to be 2
- $\lambda$: 0.5
- $\alpha$: 2 ($\gamma$ in following function represent the shape parameter of gompertz)

```{r}
#write a fn to simulate gompertz data
sim_gompertz <- function(k, gamma, N){
  #generate gompertz data
  covs <- data.frame(id = 1:N,
                    trt = stats::rbinom(N, 1, 0.5))
  dat <- simsurv(dist = "gompertz",
                 lambdas = 0.5, 
                 gammas = gamma, 
                 betas = c(trt = 2), 
                 x = covs, 
                 maxt = 5)
  dat <- merge(covs, dat)
  #fit models
  fit.exponential <- survreg(Surv(eventtime, status) ~ trt, data = dat, dist = "exponential")
  fit.weibull <- survreg(Surv(eventtime, status) ~ trt, data = dat, dist = "weibull")
  fit.cox <- coxph(Surv(eventtime, status) ~ trt, data = dat)
  
  #extract beta
  result <- tibble(exp_beta = c(-fit.exponential$coefficients[-1]), 
                  weibull_beta = c(-fit.weibull$coefficients[-1])/fit.weibull$scale,
                  cox_beta = c(fit.cox$coefficients), 
                  dist = "gompertz",
                  beta = 2, 
                  gamma = gamma,
                  N = N)
  return(result)
  }

# Set seed for reproducibility
set.seed(2023)

# Create empty dataframe to store results
sim_gompertz_result4 <- data.frame()

#gamma=2
set.seed(2023)
for (n in c(100, 150, 200, 250, 300, 350, 400)) {
for (i in 1:1000) {
  sim_res <- sim_gompertz(gamma = 2, N = n)
  sim_gompertz_result4  <- rbind(sim_gompertz_result4 , sim_res)
}
}

gompertz_table4 <- sim_gompertz_result4  %>% 
  group_by(N) %>%
  summarize(mse_exp = mean((exp_beta-2)^2),
            mse_weibull = mean((weibull_beta-2)^2),
            mse_cox = mean((cox_beta-2)^2),
            var_exp = var(exp_beta),
            var_weibull = var(weibull_beta),
            var_cox = var(cox_beta),
            bias_exp = mean(exp_beta-2),
            bias_weibull = mean(weibull_beta-2),
            bias_cox = mean(cox_beta-2)
        
  ) 
gompertz_table4  
gompertz_table4 <- gompertz_table4 %>% 
  pivot_longer(mse_exp:bias_cox, values_to = "performance", names_to = "method") %>% separate_wider_delim(method, "_", names = c("method", "model")) %>%
  pivot_wider(names_from = method, values_from = performance)
gompertz_table4
write.csv(gompertz_table4,"table/gompertz_result_table4.csv")

## Models comparison: Bias of three models using Gompertz data 
gompertz_bias <- gompertz_table4 %>% 
  ggplot(aes(x = N, y = bias, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 7",
       title="Bias vs Sample size by 3 Survival Models") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
gompertz_bias
ggsave("results/gompertz_bias.pdf", height = 5, width = 8)
ggsave("results2/gompertz_bias.pdf", height = 5, width = 8)

## Models comparison: Variance of three models using Gompertz data 
gompertz_var <- gompertz_table4 %>% 
  ggplot(aes(x = N, y = var, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 8",
       title="Variance vs Sample size by 3 Survival Models") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
gompertz_var
ggsave("results/gompertz_var.pdf", height = 5, width = 8)
ggsave("results2/gompertz_var.pdf", height = 5, width = 8)

## Models comparison: Variance of three models using Gompertz data 
gompertz_mse <- gompertz_table4 %>% 
  ggplot(aes(x = N, y = mse, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 9",
       title="MSE vs Sample size by 3 Survival Models") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
gompertz_mse
ggsave("results/gompertz_mse.pdf", height = 5, width = 8)
ggsave("results2/gompertz_mse.pdf", height = 5, width = 8)
```


# Simulation for Exponential Baseline Hazard Function

- N: sample size 100, 150, 200, 250, 300, 350, 400
- m: simulation time 1000
- $\beta$: true treatment effect to be 2
- $\lambda$: 0.5

```{r}
#write a fn to simulate exponential data
sim_exp <- function(k, N){
  #generate exponential data
  covs <- data.frame(id = 1:N,
                    trt = stats::rbinom(N, 1, 0.5))
  dat <- simsurv(dist = "exponential",
                 lambdas = 0.5, 
                 betas = c(trt = 2), 
                 x = covs, 
                 maxt = 5)
  dat <- merge(covs, dat)
  #fit models
  fit.exponential <- survreg(Surv(eventtime, status) ~ trt, data = dat, dist = "exponential")
  fit.weibull <- survreg(Surv(eventtime, status) ~ trt, data = dat, dist = "weibull")
  fit.cox <- coxph(Surv(eventtime, status) ~ trt, data = dat)
  
  #extract beta
  result <- tibble(exp_beta = c(-fit.exponential$coefficients[-1]), 
                  weibull_beta = c(-fit.weibull$coefficients[-1])/fit.weibull$scale,
                  cox_beta = c(fit.cox$coefficients), 
                  dist = "exponential",
                  beta = 2, 
                  gamma = "default",
                  N = N)
  return(result)
  }

# Set seed for reproducibility
set.seed(2023)

# Create empty dataframe to store results
sim_exp_result1  <- data.frame()

# Simulate 1000 times
for (n in c(100, 150, 200, 250, 300, 350, 400)) {
for (i in 1:1000) {
  sim_res <- sim_exp(N = n)
  sim_exp_result1 <- rbind(sim_exp_result1, sim_res)
}
}

exponential_table <- sim_exp_result1 %>% 
  group_by(N) %>%
  summarize(mse_exp = mean((exp_beta-2)^2),
            mse_weibull = mean((weibull_beta-2)^2),
            mse_cox = mean((cox_beta-2)^2),
            var_exp = var(exp_beta),
            var_weibull = var(weibull_beta),
            var_cox = var(cox_beta),
            bias_exp = mean(exp_beta-2),
            bias_weibull = mean(weibull_beta-2),
            bias_cox = mean(cox_beta-2)
        
  )

exponential_table

exponential_table <- exponential_table %>% 
  pivot_longer(mse_exp:bias_cox, values_to = "performance", names_to = "method") %>% separate_wider_delim(method, "_", names = c("method", "model")) %>%
  pivot_wider(names_from = method, values_from = performance)
exponential_table
write.csv(exponential_table,"table/exponential_result_table.csv")

## Models comparison: Variance of three models using exponential data 
exponential_bias <- exponential_table %>% 
  ggplot(aes(x = N, y = bias, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 1",
       title="Bias vs Sample size by 3 Survival Model") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
exponential_bias
ggsave("results2/exponential_bias.pdf", height = 5, width = 8)

## Models comparison: Variance of three models using exponential data 
exponential_var <- exponential_table %>% 
  ggplot(aes(x = N, y = var, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 2",
       title="Variance vs Sample size by 3 Survival Model") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
exponential_var
ggsave("results2/exponential_var.pdf", height = 5, width = 8)

## Models comparison: MSE of three models using exponential data 
exponential_mse <- exponential_table %>% 
  ggplot(aes(x = N, y = mse, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 3",
       title="MSE vs Sample size by 3 Survival Model") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
exponential_mse
ggsave("results2/exponential_mse.pdf", height = 5, width = 8)
```


# Simulation for Weibull Baseline Hazard Function

- N: sample size 100, 150, 200, 250, 300, 350, 400
- m: simulation time 1000
- $\beta$: true treatment effect to be 2
- $\lambda$: 0.5
- $\gamma = \alpha$: 2

```{r}
#write a fn to simulate weibull data
sim_weibull <- function(k, gamma, N){
  #generate weibull data
  covs <- data.frame(id = 1:N,
                    trt = stats::rbinom(N, 1, 0.5))
  dat <- simsurv(dist = "weibull",
                 lambdas = 0.5, 
                 gammas = gamma, 
                 betas = c(trt = 2), 
                 x = covs, 
                 maxt = 5)
  dat <- merge(covs, dat)
  #fit models
  fit.exponential <- survreg(Surv(eventtime, status) ~ trt, data = dat, dist = "exponential")
  fit.weibull <- survreg(Surv(eventtime, status) ~ trt, data = dat, dist = "weibull")
  fit.cox <- coxph(Surv(eventtime, status) ~ trt, data = dat)
  
  #extract beta
  result <- tibble(exp_beta = c(-fit.exponential$coefficients[-1]), 
                  weibull_beta = c(-fit.weibull$coefficients[-1])/fit.weibull$scale,
                  cox_beta = c(fit.cox$coefficients), 
                  dist = "weibull",
                  beta = 2, 
                  gamma = gamma,
                  N = N)
  return(result)
  }

# Set seed for reproducibility
set.seed(2023)

# Create empty dataframe to store results
sim_weibull_result4 <- data.frame()

# gamma=2
for (n in c(100, 150, 200, 250, 300, 350, 400)) {
for (i in 1:1000) {
  sim_res <- sim_weibull(gamma = 2, N = n)
  sim_weibull_result4 <- rbind(sim_weibull_result4, sim_res)
}
}

weibull_table4 <- sim_weibull_result4 %>% 
  group_by(N) %>%
  summarize(mse_exp = mean((exp_beta-2)^2),
            mse_weibull = mean((weibull_beta-2)^2),
            mse_cox = mean((cox_beta-2)^2),
            var_exp = var(exp_beta),
            var_weibull = var(weibull_beta),
            var_cox = var(cox_beta),
            bias_exp = mean(exp_beta-2),
            bias_weibull = mean(weibull_beta-2),
            bias_cox = mean(cox_beta-2)
        
  )
weibull_table4 <- weibull_table4 %>% 
  pivot_longer(mse_exp:bias_cox, values_to = "performance", names_to = "method") %>% separate_wider_delim(method, "_", names = c("method", "model")) %>%
  pivot_wider(names_from = method, values_from = performance)
weibull_table4
write.csv(weibull_table4,"table/weibull_result_table4.csv")


## Models comparison: Bias of three models using weibull data 
weibull_bias <- weibull_table4 %>% 
  ggplot(aes(x = N, y = bias, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 4",
       title="Bias vs Sample size by 3 Survival Model") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
weibull_bias
ggsave("results2/weibull_bias.pdf", height = 5, width = 8)

## Models comparison: Variance of three models using weibull data 
weibull_var <- weibull_table4 %>% 
  ggplot(aes(x = N, y = var, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 5",
       title="Variance vs Sample size by 3 Survival Model") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
weibull_var
ggsave("results2/weibull_var.pdf", height = 5, width = 8)

## Models comparison: MSE of three models using weibull data 
weibull_mse <- weibull_table4 %>% 
  ggplot(aes(x = N, y = mse, color = model)) +
  geom_point(size = 1) +
  geom_line(size = 1) +
  xlab("sample size") +
  labs(caption = "Figure 6",
       title="MSE vs Sample size by 3 Survival Model") + 
  theme(plot.caption = element_text(hjust = 0.5, size = rel(1.2)))
weibull_mse
ggsave("results2/weibull_mse.pdf", height = 5, width = 8)
```




# Comparing the accuracy and efficiency of treatment effect ($\beta$)

*k*: the number of independent data sets\


#### Bias

*Exponential Proportional-Hazards Model*: \
$$Bias=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{exp}^{(k)}}-\beta_{1})=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{exp}^{(k)}}-0.5)$$\

*Weibull Proportional-Hazards Model*: \
$$Bias=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{weibull}^{(k)}}-\beta_{1})=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{weibull}^{(k)}}-0.5)$$\

*Cox Proportional-Hazards Model*: \
$$Bias=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{cox}^{(k)}}-\beta_{1})=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{cox}^{(k)}}-0.5)$$\


#### Variance

*Exponential Proportional-Hazards Model*: \
$$Variance= \frac{1}{k-1}\sum_{i=1}^k\hat{(\beta_{exp}^{(k)}}-\beta_{1})^2=\frac{1}{k-1}\sum_{i=1}^k\hat{(\beta_{exp}^{(k)}}-0.5)^2$$\

*Weibull Proportional-Hazards Model*: \
$$Variance= \frac{1}{k-1}\sum_{i=1}^k\hat{(\beta_{weibull}^{(k)}}-\beta_{1})^2=\frac{1}{k-1}\sum_{i=1}^k\hat{(\beta_{weibull}^{(k)}}-0.5)^2$$\

*Cox Proportional-Hazards Model*: \
$$Variance= \frac{1}{k-1}\sum_{i=1}^k\hat{(\beta_{cox}^{(k)}}-\beta_{1})^2=\frac{1}{k-1}\sum_{i=1}^k\hat{(\beta_{cox}^{(k)}}-0.5)^2$$\


#### MSE

*Exponential Proportional-Hazards Model*: \
$$MSE= \frac{1}{k}\sum_{i=1}^k(\hat{\beta_{exp}^{(k)}}-\beta_{1})^2=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{exp}^{(k)}}-0.5)^2$$\

*Weibull Proportional-Hazards Model*: \
$$MSE= \frac{1}{k}\sum_{i=1}^k(\hat{\beta_{weibull}^{(k)}}-\beta_{1})^2=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{weibull}^{(k)}}-0.5)^2$$\

*Cox Proportional-Hazards Model*: \
$$MSE= \frac{1}{k}\sum_{i=1}^k(\hat{\beta_{cox}^{(k)}}-\beta_{1})^2=\frac{1}{k}\sum_{i=1}^k(\hat{\beta_{cox}^{(k)}}-0.5)^2$$\



#### Confidence Interval

$$Coverage: \frac{1}{k}\sum_{i=1}^kI\{\beta\in CI^{(k)}\}$$

